---
title: 'Week 4: Endogeneity, Training Exercises'
author: "Anthony Nguyen"
subtitle: <h3>Coursera/Erasmus U., Econometric Methods and Applications</h3>
output:
  pdf_document: default
  html_document:
    theme: cosmo
---
```{r load packages, include=FALSE}
library(tidyverse)
library(readxl)
library(stargazer)
library(cowplot)
library(car)
```

```{r load data for exercise, include=FALSE}

TrainExer42 <- read.csv("../../data/TrainExer42.txt")
```

\makeatletter
\g@addto@macro\bfseries{\boldmath}
\makeatother

#Training Exercise 4.2  

##Notes:

* This exercise uses the datafile `TrainExer42`and requires a computer.  
* The dataset `TrainExer42` is [available on the website](https://d396qusza40orc.cloudfront.net/eureconometrics-assets/Dataset%20Files%20for%20On-Demand%20Course/Exercises%20and%20datasets/Module%204/TrainExer42.txt).  
  
##Questions  
  
In this exercise we reconsider the example from lecture 4.1 where an analyst models sales of ice cream over time as a function of price and where price is possibly endogenous due to strategic behavior of the salesperson. In this case the salesperson knows that when a particular event is organized, demand tends to be high. Therefore she may set a high price when there is such an event.  
  
We consider the following data generating process
$$
\text{Sales} = 100 - 1 \cdot \text{Price} + \alpha \text{Event} + \varepsilon_1 \\
\text{Price} = 5 + \beta \text{Event} + \varepsilon_2
$$
where Event is a 0/1 dummy variable indicating whether an event took place at a point in time. However, when trying to estimate the price coefficient the analyst does not have the Event dummy variable and simply regresses Sales on a constant and Price.  
  
The dataset `TrainExer42` contains sales and price data for different values of $\alpha$ and $\beta$. For each scenario the same simulated values for $\varepsilon_1$ and $\varepsilon_2$ were used. Specifically, the data contains 4 price series and 16 sales series. Price variables "PriceB" give the price assuming that $\beta = B$ for $B = 0, 1, 5, 10$. Sales variables "SalesA_B" give the sales for $\alpha = A$ and $\beta = B$, where $A$ also takes the values $0, 1, 5, 10$.  
  
(a) First consider the case where the event only directly affects price ($\alpha = 0$). Estimate and report the price coefficients under all 4 scenarios for $\beta$ and calculate the $R^2$ for all these regressions. Do the estimated price coefficients signal any endogeneity problem for these values of $\alpha$ and $\beta$? Can you also explain the pattern you find for the $R^2$.  
  
(b) Repeat the exercise above, but now consider the case where the event only directly affects sales, that is, set $\beta = 0$ and check the results for the four different values of $\alpha$.  
  
(c) Finally consider the parameter estimates for the cases where the event affects price *and* sales, that is, look at $\alpha = \beta = 0, 1, 5, 10$.  Can you see the impact of endogeneity in this case?  
  
***
\pagebreak  
  
##Answers
  
### (a) First consider the case where the event only directly affects price ($\alpha = 0$). Estimate and report the price coefficients under all 4 scenarios for $\beta$ and calculate the $R^2$ for all these regressions. Do the estimated price coefficients signal any endogeneity problem for these values of $\alpha$ and $\beta$? Can you also explain the pattern you find for the $R^2$.  
  
```{r regress SALES0_ by PRICE_ , echo = TRUE, results = "asis"}

mod1a <- lm(SALES0_0~PRICE0, data = TrainExer42)
mod2a <- lm(SALES0_1~PRICE1, data = TrainExer42)
mod3a <- lm(SALES0_5~PRICE5, data = TrainExer42)
mod4a <- lm(SALES0_10~PRICE10, data = TrainExer42)

stargazer(mod1a, mod2a, mod3a, mod4a, header = FALSE, keep.stat = "rsq")  
```
When $\alpha$ is set to zero, the regression coefficients are all close to the true value of $-1$, so price is not endogenous, as the event does not influence sales directly.  
  
The $R^2$ increases for higher values of $\beta$.  This is due to the fact that for higher $\beta$ values, more of the variation in sales can be explained. In other words, for higher $\beta$ values, the variation in sales increases, and this increase is perfectly explained by the price.  
  
### (b) Repeat the exercise above, but now consider the case where the event only directly affects sales, that is, set $\beta = 0$ and check the results for the four different values of $\alpha$.     
  
```{r regress , echo = TRUE, results = "asis"}

mod1b <- lm(SALES0_0~PRICE0, data = TrainExer42)
mod2b <- lm(SALES1_0~PRICE0, data = TrainExer42)
mod3b <- lm(SALES5_0~PRICE0, data = TrainExer42)
mod4b <- lm(SALES10_0~PRICE0, data = TrainExer42)

stargazer(mod1b, mod2b, mod3b, mod4b, header = FALSE, keep.stat = "rsq")  
```
Again, here we can see that all of the regression coefficients are close to the true value of $-1$, thus we can say that Price is not endogenous, as the Event only affects Sales and not Price.  
  
Therefore, the omission of the EVENT variable, does not lead to a correlation between the error term and Price.  
  
From the regression results, we can also see that the $R^2$ term drops significantly for higher values of $\alpha$. At a high value of $\alpha$, a lot of variation in Sales is due to the Event, however, this variation is not captured in the regression because we only regressed Sales on a Constant and Price. This is also the reason that the estimate for $\alpha$ as $10$ is relatively small ($0.08$), as it reflects the relatively large estimation uncertainty.  
  
### (c) Finally consider the parameter estimates for the cases where the event affects price *and* sales, that is, look at $\alpha = \beta = 0, 1, 5, 10$.  Can you see the impact of endogeneity in this case?  
  
```{r regress, echo = TRUE, results = "asis"}

mod1c <- lm(SALES0_0~PRICE0, data = TrainExer42)
mod2c <- lm(SALES1_1~PRICE1, data = TrainExer42)
mod3c <- lm(SALES5_5~PRICE5, data = TrainExer42)
mod4c <- lm(SALES10_10~PRICE10, data = TrainExer42)

stargazer(mod1c, mod2c, mod3c, mod4c, header = FALSE, keep.stat = "rsq")    
```
Here, on the diagonal of the regression summary, we *can* see consequences of endogeneity.  If $\alpha$ and $\beta$ are both non-zero, the omission of the EVENT dummy will lead to correlation between the error term in the regression in PRICE.  
As a consequence of this correlation, the estimate can be completely off.  For instance, we can see in the case where $\alpha$ and $\beta$ is equal to $10$ the estimate is almost zero ($-0.09$).
      
      
    
