---
title: 'Week 1: Simple Regression, Training Exercise 1.2'
author: "Anthony Nguyen"
subtitle: <h3>Coursera/Erasumus U., Econometric Methods and Applications</h3>
output:
  html_document:
    theme: cosmo
---
```{r load packages, include=FALSE}
library(tidyverse)
```

```{r load data for exercise, include=FALSE}
TrainExer11 <- read_tsv("../../data/TrainExer11.txt")
```

#Questions  
A transformation of the data on $x_i$ and  $y_i$ (like taking their logarithm) changes the interpretation of the slope parameter $\beta$.  
 
  (a) Show that in the regression model $log(yi) = \alpha + \beta log(x_i) + \varepsilon_ i$, the elasticity of $y$ with respect to $x$ is equal to $\beta$ (that is, does not depend on the values of xi and yi).  
  
  (b) Determine the elasticity of $y$ with respect to $x$ in the model $y_i = \alpha + \beta log(x_i) + \varepsilon_i$.  
    
  (c) Determine the elasticity of $y$ with respect to $x$ in the model $log(yi) = \alpha + \beta x_i + \varepsilon_i$.  
    
***  

#Answers  
  
(a) 

$$
\begin{equation} \label{eq1}
\begin{split}
A & = \frac{\pi r^2}{2} \\
 & = \frac{1}{2} \pi r^2
\end{split}
\end{equation}
$$
