---
title: 'Week 1: Simple Regression, Training Exercise 1.1'
author: "Anthony Nguyen"
subtitle: <h3>Coursera/Erasumus U., Econometric Methods and Applications</h3>
output:
  html_document:
    keep_md: yes
    theme: cosmo
---
```{r load packages, include=FALSE}
library(tidyverse)
```

```{r load data for exercise, include=FALSE}
TrainExer11 <- read_tsv("../../data/TrainExer11.txt")
```

#Questions  

Dataset `TrainExer11`contains survey outcomes of a travel agency that wishes to improve recommendation strategies for its clients. The dataset contains 26 observations on age and average daily expenditures during holidays.
  
  (a) Make two histograms,one of expenditures and the other of age. Make also a scatterdiagram with expenditures on the vertical axis versus age on the horizontal axis.  
  
  (b) In what respect do the data in this scatter diagram look different from the case of the sales and price data discussed in the lecture?  
  
  (c) Propose a method to analyze these data in a way that assists the travel agent in making recommendations to future clients.   
  
The scatter diagram indicates two groups of clients. Younger clients spend more than older ones. Further, expenditures tend to increase with age for younger clients, whereas the pattern is less clear for older clients.  
  
  (d) Compute the sample mean of expenditures of all 26 clients.    
  
  (e) Compute two sample means of expenditures, one for clients of age forty or more and the other for clients of age below forty.    
  
  (f) What daily expenditures would you predict for a new client of fifty years old? And for someone who is twenty-five years old?  
    
***    
    
#Answers  
  
(a) Summary of data:

```{r quick summary of data, echo = FALSE}
summary(TrainExer11[, 2:3])
```

```{r, plot histogram of expeditures, echo = FALSE}
hist_expenditures <- TrainExer11 %>% ggplot(aes(x = Expenditures)) +
  geom_histogram(binwidth = 2) 
hist_expenditures
```
```{r, plot histogram of age, echo = FALSE}
hist_age <- TrainExer11 %>% ggplot(aes(x=Age)) +
  geom_histogram(binwidth = 2)
hist_age
```
```{r, plot scatterplot of age v expenditure, echo = FALSE}
age_expenditure <- TrainExer11 %>% ggplot(aes(x = Age, y = Expenditures)) +
  geom_point()
age_expenditure
```
  
(b)  The results of this scatter diagram look quite different from the example in the lecture video, primarily because the data here is not all centered around one easily identifiable line.  In fact, it looks like the data is clustered into two distinct groups here, one witt a fairly linear looking relationship (age < 40) and another group that does not seem to have a strong linear relationship (age > 40).  
  
(c)  Based on the data presented here, I would split the sample into two groups (under and over 40) and build my reccomendations to the travel agent that way.  In general, the younger group spends more money, and that spending tends to rise as they get closer to 40.  For the older group, their average spending looks much lower, and the relationship between age and expense is less clear.  
  
(d)

```{r compute mean expediture for all, echo = FALSE}

mean(TrainExer11$Expenditures)

```

(e)  
  
```{r, compute mean expenditure for under and over 40s, echo = FALSE}
under40 <- TrainExer11 %>% filter(Age < 40) 
mean(under40$Expenditures)

over40 <- TrainExer11 %>% filter(Age >= 40)
mean(over40$Expenditures)

```
  
(f)  for a new 50 year old client, I would expect an average expenditure of around 95.8, while the 25 year old should be around 106.4.
  



