---
title: 'Week 3: Model Specification, Training Exercises'
author: "Anthony Nguyen"
subtitle: <h3>Coursera/Erasmus U., Econometric Methods and Applications</h3>
output:
  pdf_document: default
  html_document:
    theme: cosmo
---
```{r load packages, include=FALSE}
library(tidyverse)
library(readxl)
library(stargazer)
library(cowplot)
```

```{r load data for exercise, include=FALSE}
TrainExer31 <- read_excel("../../data/TrainExer 3-1.xlsx")

```

\makeatletter
\g@addto@macro\bfseries{\boldmath}
\makeatother

#Training Exercise 3.1  

##Notes:

* This exercise uses the datafile `TrainExer31`and requires a computer.  
* The dataset `TrainExer31` is [available on the website](https://d396qusza40orc.cloudfront.net/eureconometrics-assets/Dataset%20Files%20for%20On-Demand%20Course/Exercises%20and%20datasets/Module%203/TrainExer%203-1-corrected.txt).  
  
##Questions  
  
(a) Use dataset `TrainExer31` to regress the change in the log of the S&P500 index on a constant and the book-to-market ration, and check the result presented in Lecture 3.1 that:  
$$
\text{change in log(SP500 index)}= 0.177 - 0.213 \cdot \text{Book-to-market} + e
$$

(b) Now regress the S&P500 index (without any kind of transformation) on a constant and the book-to-market ratio.  Consider whether the effect of the book-to-market on the index is significant in this specification.  
    
(c) Make a plot of the residuals $e$ from both question (a) and (b) and comment on the difference. 
  
***
\pagebreak  
  
##Answers

### (a) Use dataset `TrainExer31` to regress the change in the log of the S&P500 index on a constant and the book-to-market ratio, and check the result presented in Lecture 3.1 that:  
$$
\text{change in log(SP500 index)}= 0.177 - 0.213 \cdot \text{Book-to-market} + e
$$

We need to be careful here, as the question is asking for the *change* in the log of the S&P500 index, not just the log. In `R` we can do this with the `diff()` function, being careful to pad by one row so that our vector is the same length as the rest of the dataframe:
```{r take diff log of Index, echo=TRUE}

#take log of Index variable
TrainExer31 <- TrainExer31 %>% mutate(log_Index = log(Index))  
  
#take difference (change) of log of Index, adding NA to pad first row  
TrainExer31 <- TrainExer31 %>% mutate(diff_log_Index = c(NA, diff(log_Index, lag=1)))
```

Now that we've calculated the change in the log of the Index, we can run the regression with our transformed y-variable:
```{r regress log_Index~BookMarket, echo = TRUE}

mod_difflog_Index_BookMarket <- lm(diff_log_Index~BookMarket, data = TrainExer31)

summary(mod_difflog_Index_BookMarket)
```
  
### (b) Now regress the S&P500 index (without any kind of transformation) on a constant and the book-to-market ratio. Consider whether the effect of the book-to-market on the index is significant in this specification.  

```{r regress Index~BookMarket, echo = TRUE}

mod_Index_BookMarket <- lm(Index~BookMarket, data = TrainExer31)

summary(mod_Index_BookMarket)
```
Looking at the results of this regression, the effect of BookMarket on Index has a t-value of -8.075, and p-value that is significant at the 99% level.  
  
### (c) Make a plot of the residuals $e$ from both question (a) and (b) and comment on the difference.  

```{r plot resideuals from both models, echo = TRUE}
res_a <- qplot(mod_difflog_Index_BookMarket$residuals, xlab = "e, diff_log_Index~BookMarket")
res_b <- qplot(mod_Index_BookMarket$residuals, xlab = "e, Index~BookMarket") 

plot_grid(res_a, res_b)
```  
From the plot we can see that the distribution of the residuals from the two regressions have a very different distribution and use different scales. 
  

